Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-09-20 manuscript/paper.tex
longer equations inside widetext -some minor changes

We consider
rigid rotation and tilting of the 
%BO$_{6/2}$ 
octahedra
 in ABO$_3$ perovskites
 from the viewpoint of Jahn-Teller problem
 and study the
effect of the crystal field of the nearest A and B ion cages 
%that immediately surround the octahedra
on the single electron spectrum of the 
central 
transition metal ions.
While the crystal field of the octahedra 
that 
% quenches the orbital angular momentum of the central B ion and 
creates $e_g$ and $t_{2g}$ manifolds
do not change with such rigid rotations, 
the field of 
%A and B ionic cages %that immediately surround the octahedra
A and B cages
deforms significantly %with the octahedral rotations 
from the ideal octahedral field
in a way that
leads to sizeable splittings in the $e_g$ and $t_{2g}$ manifolds of the central B ion.
%Similar to the Jahn-Teller problem,
The lowering of the ground state energy
of a given many-electron configuration
due to these splittings
can thus be an important driving force behind the octahedral rotations 
and associated structural transformations.
We find that the size of the splitting and the local orbital ordering are
determined by a competition between the A and B cages, 
as well as a competition between the tilt and rotation angles.

@@

%Their crystal structures 
The perovskite structures
have a common formula unit ABX$_3$,
where 
%positive ions A and negative ions X form a cube and an octahedron around positive ion B.
A and B are positive ions while X are negative ions 
%whose six-fold coordination forms octahedra around B.
that form octahedra around  B.
In a cubic perovskite,
the octahedra are symmetric, 
but they can 
distort~\cite{
BersukerCCR75,
KugelSPU82,
MillisN98,
TerakuraPMS07,
HalcrowRSC13,
LufasoAC04}
or simply 
rotate~\cite{
GlazerAC72, ThomasAC96,
WoodwardAC97a,WoodwardAC97b,
HowardAC98,StokesAC02,AngelPRL05}
 to lower the symmetry
and produce, e.g., a rhombohedral, tetragonal or orthorhombic structure.
% cubic perovskite [spacegroup 221] can transform to:
% crystal system: spacegroups
% triclinic: 2
% monoclinic: 11,12,15
% orthorhombic: 59,62,63,71,74
% tetragonal 127,139,140
% rhombohedral/trigonal: 167
% cubic: 204, 221 
Octahedral rotations also occur in other perovskite related structures
and
have been extensively 
studied
due to their crucial role in determining various physical properties,
e.g., in 
ferroelectric~\cite{GhosezNM11,
SchickPRB14,
SimPRB14,
LeePRL17,
WangPRL22,
ZhouPRB22,
GaoPRB23} and 
multiferroic~\cite{BenedekPRL11,
YangPRL14,
YePRB18,
ZhangPRL20,
ZhangYajunPRL20, 
ZhouPRB21,
ZhangPRL22}
systems.
The interplay of octahedral rotations
and 
spin-orbit coupling~\cite{AmatNL14,KrachJPCM23},
breathing distortions~\cite{BalachandranPRB13},
strain~\cite{Aguado-PuentePRL11},
defects~\cite{JiaPRB22}, and
pressure~\cite{XiangPRB17},
and 
their role in
magnetostructural phase transitions~\cite{JohnsonPRL20},
orbital order~\cite{OgawaPRL12,ChenPRL19}, orbital anisotropy~\cite{HuangPRB21}, 
and
%surface electronic structure under in-plane uniaxial compression~\cite{MoralesPRL23}
surface electronic structure~\cite{MoralesPRL23},
has also been investigated.
% JT interacting with octa tilting:
Since, usually, the octahedra are not just %(Jahn-Teller) 
distorted or rotated but a combination of both,
the influence of these two 
structural mutations %instabilities 
on each other has recently been explored by many authors~\cite{LufasoAC04,CarpenterAC09,ChenPRL19}.

@@

Beyond the anion octahedron,
each transition metal ion B in a cubic perovskite
is coordinated by two cation cages,
a cube of A and an octahedron of B, as shown in Figs.~\ref{fig:struc}(c,d).
Including the 
%electrostatic 
potential of these two ion cages
into the 
crystal field Hamiltonian of 
the central ion B,
we explore the effect 
of the rotation and tilting on 
the splittings in its $t_{2g}$ and $e_g$ manifolds
that are
created primarily by the anion's octahedral potential.
We find that the size of the splitting 
and local orbital ordering
depend
on the relative charges on the two types of cations
and a competition between rotation and tilting.
At typical parameter values,
the splitting is large enough 
to play a major role in 
the corresponding structural instability. 
The qualitative picture developed by our calculations 
holds even when the potential of the full lattice is considered.

@@

The paper is organised as follows.
Section~\ref{sec:formalistm}
presents the formalism and 
sec.~\ref{sec:JT} reviews the Jahn-Teller effect to
introduce the notation and concepts involved
by considering tetragonal distortion of an octahedron.
%\az{[.... multiple subsections?]}
The model and calculations are presented
in sec.~\ref{sec:calc},
where
we first discuss the relative contributions of the three closest ion cages,
in the cubic structure, to
the crystal field at the central transition metal site 
and crystal field splitting between its $e_g$ and $t_{2g}$ manifolds
(sec.~\ref{sec:relativeVoctABO}),
and then 
describe the deformation effects of 
rigid octahedral rotations (sec.~\ref{sec:rotation})
on
the structure (sec.~\ref{sec:structure}) 
and crystal field (sec.~\ref{sec:deformCF}).
The results are in sec.~\ref{sec:results},
where we %seperately
present the splitting in the single particle spectrum
belonging to the $t_{2g}$ (sec.~\ref{sec:t2g})
and $e_g$ (sec.~\ref{sec:eg-gen}) manifolds,
and finally describe the effect of the full lattice (sec.~\ref{sec:FL}).
\az{appendix?}

@@

In a cubic perovskite ABO$_3$,
each transition metal atom/ion B
is surrounded by six O$^{2-}$, each at a distance $a/2$,
 making an octahedron of size $a$ around it.
The potential of these ions $V_{O}(\vec r)= V_{oct}(\vec r)$
splits the degenerate d-orbitals
by an amount ${\Delta_O=160q_oD_{4}/3a}$
 into $t_{2g}$ and $e_g$.
However, this is not the only crystal field potential at the B site.
The coordination cages of next nearest neighbours, 
Eight A atoms/cations all at the same distance $\sqrt{3}a/2$
make a cube around the B ion,
which also produces an octahedral field, given by 
\begin{align}
V_{A}(\vec r) = \frac{8}{81 \sqrt{3}}\frac{q_A}{q_o} V_{O}(\vec r).
\end{align}
Furthermore,
the nearest neighbouring B ions, size of them each at a distance $a$,
make an octahedron, with a potential 
\begin{align}
V_{B}(\vec r) = -\frac{1}{32}\frac{q_B}{q_o}V_{O}(\vec r).
\end{align}
It is interesting to note that 
$V_{A}(\vec r)$ and $V_{B}(\vec r)$ have opposite signs
and 
their combined effect 
$V_{AB}(\vec r) = V_{A}(\vec r)+ V_{B}(\vec r)$
depends on their relative charges.
Using the charge neutrality $q_B=3q_o-q_A$,
we can write 
\begin{align}
V_{AB}(\vec r) &= f_{q_A}V_{O}(\vec r),\\
f_{q_A} &= \left(\frac{1}{32}+\frac{8}{81 \sqrt{3}}\right)\frac{q_A}{q_o} -\frac{3}{32},
\end{align}
which shows that $V_{AB}(\vec r)\to 0$
as
${q_A/q_0\to 729/({256 \sqrt{3}+243})\simeq 1.062}$.
%where they cancel each other exactly.

@@

@@

V^A_{4,-2}&= \frac{128}{3} \sqrt{5 \pi } \cos ^2\beta  \left(\frac{3 \cos 2 \alpha -\cos 2 \beta +2}{\left(\cos ^2\alpha +2 \cos ^2\beta \right)^{9/2}}+\frac{\sec ^5\alpha  (\cos 2 \beta -5)}{(\cos 2 \beta +2)^{9/2}}\right),\\ \nonumber
V^A_{4,0}&= -\frac{16}{3} \sqrt{\pi } \left(\frac{8 \cos 2 \alpha  (3 \cos 2 \beta +2)-2 \cos 4 \alpha+12 \cos 2 \beta -3 \cos 4 \beta +9}{\left(\cos ^2\alpha +2 \cos ^2\beta \right)^{9/2}}\right. \\
&+\left. \frac{\sec ^5\alpha  (36 \cos 2 \beta -3 \cos 4 \beta +23)}{(\cos 2 \beta +2)^{9/2}}\right),\\
V^A_{4,4}&=-\frac{128}{3} \sqrt{35 \pi } \cos ^4\beta  \left(\frac{1}{\left(\cos ^2\alpha +2 \cos ^2\beta \right)^{9/2}}+\frac{\sec ^5\alpha }{(\cos 2 \beta +2)^{9/2}}\right),

@@

@@

V^{B}_{2,-2}&=  \frac{64 \sqrt{\frac{3 \pi }{5}} \sin ^2\alpha  \sec ^3\beta }{(\cos 2 \alpha +3)^{5/2}},

@@

 V^{B}_{4,-2}&=  -\frac{128 \sqrt{5 \pi } \sin ^2\alpha  \sec ^5\beta }{3 (\cos 2 \alpha +3)^{7/2}},\\ 
V^{B}_{4,0}&= \frac{4}{3} \sqrt{\pi } \left(\frac{24 \sec ^5\beta }{(\cos 2 \alpha +3)^{5/2}}+\sec ^5\alpha \right),\\
 V^{B}_{4,4}&=  -\frac{16 \sqrt{35 \pi } (-20 \cos 2 \alpha +\cos 4 \alpha -13) \sec ^5\beta }{3 (\cos 2 \alpha +3)^{9/2}}.

@@

Ignoring the coupling between the $e_{g}$ and $t_{2g}$ manifolds (between $xy$ and $x^2-y^2$ at $\alpha=0$),
their Hamiltonians turn out to be diagonal.
We can write the single particle energies of $t_{2g}$ states
as
\begin{align}
\label{eq:theta0-Es}
(E_{xy},&E_{yz},E_{zx})= (2, -1, -1)
\left[\frac{}{} \Delta_{2}  D_2 + \Delta_4 D_4\right],
 \end{align}
 where $E_{yz}$ and $E_{zx}$ are degenerate and 
\begin{align}
\Delta_{l} =&  \left(q_A \Delta_{l}^A + q_B\Delta_{l}^B\right)/a,~(l=2,4),\\
\Delta_{2}^A =& 
%from A
\frac{64 \sin ^2\beta }{7 (\cos 2 \beta +2)^{5/2}} % simplifies.... from:
 %\frac{16 (3 \sin \beta+\sin 3\beta)^2}{7 (\cos 2\beta+2)^{9/2}},\\
 \simeq \frac{64}{63 \sqrt{3}}\beta ^2,\\
\Delta_4^A =&
 \frac{160 \sin ^2\beta}{63 (\cos 2\beta+2)^{9/2}} \times \\
&  (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82),\\
\simeq& \frac{40960}{5103 \sqrt{3}}\beta ^2,\\
% \Delta_4^A =&
% \frac{160 \sin ^2\beta (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82)}{63 (\cos 2\beta+2)^{9/2}},\\
 %from B
 \label{eq:theta0-delB2}
\Delta_{2}^B =& \frac{2}{7} \left(1-\sec ^3\beta\right)
\simeq -\frac{3}{7}\beta ^2,\\
\label{eq:theta0-delB4}
  \Delta_4^B =&\frac{5}{126} \left((7 \cos 4\beta-3) \sec ^5\beta-4\right)
  \simeq -\frac{115}{63}\beta ^2.
 %\left( \Delta_{2}  D_2 + \Delta_4 D_4\right),\\
 \end{align}
 We see that $\Delta_{2/4}^A>0$ while $\Delta_{2/4}^B<0$,
 revealing a competition between the two terms 
 that would determine the sign and size of the term
  in the square bracket in Eq.~\ref{eq:theta0-Es} 
%  (hence the ground state and splitting) 
  as $q_A$ (and hence $q_B$)
  is varied.
The average $t_{2g}$ energy
 also shifts from
$-2/5\Delta$ by an amount
%\rev{E0t2g or change in it? how to define the expressions... plotted is always change in it...}
\begin{align}
\label{eq:theta0-E0}
E_{0}^{t_{2g}} =& \left(q_A E_{0}^A + q_BE_{0}^B\right)D_4/a
-(2/5)f_{q_A}\Delta_{O},\\
E_{0}^A =&
 \frac{8 (92 \cos 2\beta+24 \cos 4\beta+20 \cos 6\beta+5 \cos 8\beta+51)}{9 (\cos 2\beta+2)^{9/2}},\\
 \simeq& \frac{512}{243 \sqrt{3}}-\frac{2560 \beta ^2}{729 \sqrt{3}},\\
E_{0}^B=&
 -\frac{1}{18} \left((5 \cos 4\beta+3) \sec ^5\beta+4\right)
 \simeq -\frac{2}{3}+\frac{10 \beta ^2}{9}.
 \end{align}

>>>>

We consider
rigid rotation and tilting of the 
%BO$_{6/2}$ 
octahedra
 in ABO$_3$ perovskites
 from the viewpoint of Jahn-Teller problem
 and study the
effect of the crystal field of the nearest A and B ion cages 
%that immediately surround the octahedra
on the single electron spectrum of B, the 
central 
transition metal ions.
While the crystal field of the octahedra 
that 
% quenches the orbital angular momentum of the central B ion and 
creates degenerate manifolds $e_g$ and $t_{2g}$ 
do not change with such rigid rotations, 
the field of 
%A and B ionic cages %that immediately surround the octahedra
A and B cages
deforms significantly %with the octahedral rotations 
from the ideal octahedral field
in a way that
leads to sizeable splittings in 
%the $e_g$ and $t_{2g}$ manifolds of the central B ion.
these manifolds.
%Similar to the Jahn-Teller problem,
The lowering of the ground state energy
of a given many-electron configuration
due to these splittings
can thus be an important driving force behind the octahedral rotations 
and associated structural transformations.
We find that the size of the splitting and the local orbital ordering are
determined by a competition between the A and B cages, 
as well as a competition between the tilt and rotation angles.

@@

%Their crystal structures 
The perovskite structures
have a common formula unit ABX$_3$,
where 
%positive ions A and negative ions X form a cube and an octahedron around positive ion B.
A and B are positive ions while X are negative ions 
%whose six-fold coordination forms octahedra around B.
that form octahedra around  B.
In a cubic perovskite,
the octahedra are symmetric, 
but they can 
distort~\cite{
BersukerCCR75,
KugelSPU82,
MillisN98,
TerakuraPMS07,
HalcrowRSC13,
LufasoAC04}
or simply 
rotate~\cite{
GlazerAC72, ThomasAC96,
WoodwardAC97a,WoodwardAC97b,
HowardAC98,StokesAC02,AngelPRL05}
 to lower the symmetry
and produce, e.g., a rhombohedral, tetragonal or orthorhombic structure.
% cubic perovskite [spacegroup 221] can transform to:
% crystal system: spacegroups
% triclinic: 2
% monoclinic: 11,12,15
% orthorhombic: 59,62,63,71,74
% tetragonal 127,139,140
% rhombohedral/trigonal: 167
% cubic: 204, 221 
Octahedral rotations also occur in other perovskite related structures
and
have been extensively 
studied
due to their crucial role in determining various physical properties,
e.g., in 
ferroelectric~\cite{GhosezNM11,
SchickPRB14,
SimPRB14,
LeePRL17,
WangPRL22,
ZhouPRB22,
GaoPRB23} and 
multiferroic~\cite{BenedekPRL11,
YangPRL14,
YePRB18,
ZhangPRL20,
ZhangYajunPRL20, 
ZhouPRB21,
ZhangPRL22}
systems.
The interplay of octahedral rotations
and 
spin-orbit coupling~\cite{AmatNL14,KrachJPCM23},
breathing distortions~\cite{BalachandranPRB13},
strain~\cite{Aguado-PuentePRL11},
defects~\cite{JiaPRB22}, and
pressure~\cite{XiangPRB17},
and 
their role in
magnetostructural phase transitions~\cite{JohnsonPRL20},
orbital order~\cite{OgawaPRL12,ChenPRL19}, orbital anisotropy~\cite{HuangPRB21}, 
and
%surface electronic structure under in-plane uniaxial compression~\cite{MoralesPRL23}
surface electronic structure~\cite{MoralesPRL23},
has also been investigated.
% JT interacting with octa tilting:
Since, usually, the octahedra are not just %(Jahn-Teller) 
distorted or rotated but a combination of both,
the influence of these two 
structural mutations %instabilities 
on each other has recently been explored by many authors~\cite{LufasoAC04,CarpenterAC09,ChenPRL19}
as well.

@@

Beyond the anion octahedron,
each transition metal ion B in a cubic perovskite
is coordinated by two cation cages,
a cube of A and an octahedron of B, as shown in Figs.~\ref{fig:struc}(c,d).
Including the 
%electrostatic 
potential of these two ion cages
into the 
crystal field Hamiltonian of 
the central ion B,
we explore the effect 
of the rotation and tilting on 
%the splittings in 
its $t_{2g}$ and $e_g$ manifolds
that are
created primarily by the anion's octahedral potential.
We find that the degeneracy of $t_{2g}$ and $e_g$
is lifted, and
the size of the splitting 
and local orbital ordering
depend
on the relative charges on the two types of cations
and a competition between rotation and tilting.
At typical parameter values,
the splitting is large enough 
to play a major role in 
the corresponding structural instability. 
The qualitative picture developed by our calculations 
holds even when the potential of the full lattice is considered.

@@

The paper is organised as follows.
Section~\ref{sec:formalistm}
presents the formalism and 
sec.~\ref{sec:JT} reviews the Jahn-Teller effect to
introduce the notation and concepts involved
by considering tetragonal distortion of an octahedron.
%\az{[.... multiple subsections?]}
The model and calculations are presented
in sec.~\ref{sec:calc},
where
we first discuss the relative contributions of the three closest ion cages
in the cubic structure
 to
the crystal field at the central transition metal site 
and crystal field splitting between its $e_g$ and $t_{2g}$ manifolds
(sec.~\ref{sec:relativeVoctABO}),
and then 
describe the deformation effects of 
rigid octahedral rotations (sec.~\ref{sec:rotation})
on
the structure (sec.~\ref{sec:structure}) 
and crystal field (sec.~\ref{sec:deformCF}).
The results are in sec.~\ref{sec:results},
where we %seperately
present the splitting in the single particle spectrum
belonging to the $t_{2g}$ (sec.~\ref{sec:t2g})
and $e_g$ (sec.~\ref{sec:eg-gen}) manifolds,
and finally describe the effect of the full lattice (sec.~\ref{sec:FL}).
\az{appendix?}

@@

In a cubic perovskite ABO$_3$,
each transition metal atom/ion B
is surrounded by six O$^{2-}$, each at a distance $a/2$,
 making an octahedron of size $a$ around it.
The potential of these ions $V_{O}(\vec r)= V_{oct}(\vec r)$
splits the degenerate d-orbitals
by an amount ${\Delta_O=160q_oD_{4}/3a}$
 into $t_{2g}$ and $e_g$.
However, this is not the only crystal field potential at the B site.
The coordination cages of next nearest neighbours, 
Eight A atoms/cations all at the same distance $\sqrt{3}a/2$
make a cube around the B ion,
which also produces an octahedral field, given by 
\begin{align}
V_{A}(\vec r) = \frac{8}{81 \sqrt{3}}\frac{q_A}{q_o} V_{O}(\vec r).
\end{align}
Furthermore,
the nearest neighbouring B ions, size of them each at a distance $a$,
make an octahedron, with a potential 
\begin{align}
V_{B}(\vec r) = -\frac{1}{32}\frac{q_B}{q_o}V_{O}(\vec r).
\end{align}
It is interesting to note that 
$V_{A}(\vec r)$ and $V_{B}(\vec r)$ have opposite signs
and 
their combined effect 
$V_{AB}(\vec r) = V_{A}(\vec r)+ V_{B}(\vec r)$
depends on their relative charges.
Using the charge neutrality $q_B=3q_o-q_A$,
we can write 
\begin{align}
V_{AB}(\vec r) &= f_{q_A}V_{O}(\vec r),\\
f_{q_A} &= \left(\frac{1}{32}+\frac{8}{81 \sqrt{3}}\right)\frac{q_A}{q_o} -\frac{3}{32},
\end{align}
which shows that ${V_{AB}(\vec r)}$ vanishes
at
$q_A/q_o=729/({256 \sqrt{3}+243}) {\simeq 1.062}$.
%where they cancel each other exactly.

@@

\begin{widetext}

@@

V^A_{4,-2}&= \frac{128\sqrt{5 \pi } }{3} \cos ^2\beta  \left(\frac{3 \cos 2 \alpha -\cos 2 \beta +2}{\left(\cos ^2\alpha +2 \cos ^2\beta \right)^{9/2}}+\frac{\sec ^5\alpha  (\cos 2 \beta -5)}{(\cos 2 \beta +2)^{9/2}}\right),\\ \nonumber
V^A_{4,0}&= \frac{16\sqrt{\pi } }{3} \left(
\frac{\sec ^5\alpha  (36 \cos 2 \beta -3 \cos 4 \beta +23)}{(\cos 2 \beta +2)^{9/2}}
-
\frac{8 \cos 2 \alpha  (3 \cos 2 \beta +2)-2 \cos 4 \alpha+12 \cos 2 \beta -3 \cos 4 \beta +9}{\left(\cos ^2\alpha +2 \cos ^2\beta \right)^{9/2}}
\right),\\
V^A_{4,4}&=-\frac{128\sqrt{35 \pi } }{3} \cos ^4\beta  \left(\frac{1}{\left(\cos ^2\alpha +2 \cos ^2\beta \right)^{9/2}}+\frac{\sec ^5\alpha }{(\cos 2 \beta +2)^{9/2}}\right),

@@

\end{widetext}

@@

V^{B}_{2,-2}&= 64 \sqrt{\frac{3 \pi }{5}}  \frac{\sin ^2\alpha  \sec ^3\beta }{(\cos 2 \alpha +3)^{5/2}},

@@

 V^{B}_{4,-2}&=  - \frac{128 \sqrt{5 \pi } }{3} \frac{\sin ^2\alpha  \sec ^5\beta }{ (\cos 2 \alpha +3)^{7/2}},\\ 
V^{B}_{4,0}&= \frac{4\sqrt{\pi } }{3} \left(\frac{24 \sec ^5\beta }{(\cos 2 \alpha +3)^{5/2}}+\sec ^5\alpha \right),\\
 V^{B}_{4,4}&=  - \frac{16 \sqrt{35 \pi } }{3} \frac{(-20 \cos 2 \alpha +\cos 4 \alpha -13) %\sec ^5\beta }
 {(\cos 2 \alpha +3)^{9/2} \cos^5\beta}.

@@

Ignoring the coupling between the $e_{g}$ and $t_{2g}$ manifolds (between $xy$ and $x^2-y^2$ at $\alpha=0$),
their Hamiltonians turn out to be diagonal.
We can write the single particle energies of $t_{2g}$ states
as
\begin{align}
\label{eq:theta0-Es}
(E_{xy},&E_{yz},E_{zx})= (2, -1, -1)
\left[\frac{}{} \Delta_{2}  D_2 + \Delta_4 D_4\right],
 \end{align}
 where $E_{yz}$ and $E_{zx}$ are degenerate and 
\begin{align}
\Delta_{l} =&  \left(q_A \Delta_{l}^A + q_B\Delta_{l}^B\right)/a,~(l=2,4),\\
\Delta_{2}^A =& 
%from A
\frac{64 \sin ^2\beta }{7 (\cos 2 \beta +2)^{5/2}} % simplifies.... from:
 %\frac{16 (3 \sin \beta+\sin 3\beta)^2}{7 (\cos 2\beta+2)^{9/2}},\\
 \simeq \frac{64}{63 \sqrt{3}}\beta ^2,\\
 \nonumber
\Delta_4^A =&
 \frac{160 \sin ^2\beta}{63 (\cos 2\beta+2)^{9/2}} \times \\
&  (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82)\\
\simeq& \frac{40960}{5103 \sqrt{3}}\beta ^2,\\
% \Delta_4^A =&
% \frac{160 \sin ^2\beta (125 \cos 2\beta+42 \cos 4\beta+7 \cos 6\beta+82)}{63 (\cos 2\beta+2)^{9/2}},\\
 %from B
 \label{eq:theta0-delB2}
\Delta_{2}^B =& \frac{2}{7} \left(1-\sec ^3\beta\right)
\simeq -\frac{3}{7}\beta ^2,\\
\label{eq:theta0-delB4}
  \Delta_4^B =&\frac{5}{126} \left((7 \cos 4\beta-3) \sec ^5\beta-4\right)
  \simeq -\frac{115}{63}\beta ^2.
 %\left( \Delta_{2}  D_2 + \Delta_4 D_4\right),\\
 \end{align}
 We see that $\Delta_{2/4}^A>0$ while $\Delta_{2/4}^B<0$,
 revealing a competition between the two terms 
 that would determine the sign and size of the term
  in the square bracket in Eq.~\ref{eq:theta0-Es} 
%  (hence the ground state and splitting) 
  as $q_A$ (and hence $q_B$)
  is varied.
The average $t_{2g}$ energy
 also shifts from
$-2/5\Delta$ by an amount
%\rev{E0t2g or change in it? how to define the expressions... plotted is always change in it...}
\begin{align}
\label{eq:theta0-E0}
E_{0}^{t_{2g}} = \left(q_A E_{0}^A + q_BE_{0}^B\right)D_4/a
-(2/5)f_{q_A}\Delta_{O},
 \end{align}
 where
\begin{align}
\nonumber
E_{0}^A =&
 \frac{8 (92 \cos 2\beta+24 \cos 4\beta+20 \cos 6\beta+5 \cos 8\beta+51)}{9 (\cos 2\beta+2)^{9/2}}\\
 \simeq& \frac{512}{243 \sqrt{3}}-\frac{2560 \beta ^2}{729 \sqrt{3}},\\
 \nonumber
E_{0}^B=&
 -\frac{1}{18} \left((5 \cos 4\beta+3) \sec ^5\beta+4\right)\\
 \simeq& -\frac{2}{3}+\frac{10 \beta ^2}{9}.
 \end{align}
